Design and Implementation of Probabilistic Programming Language Anglican

### 2016-11-27 paper/paper.tex
update

The simplest inference algorithm is \textit{importance
sampling}:
\begin{lstlisting}[style=default]
(derive ::algorithm
        :anglican.inference/algorithm)
(defmethod infer :importance [_ prog value & {}]
  (letfn
    [(sample-seq []
       (lazy-seq
         (cons
           (:state (exec ::algorithm prog value
                         initial-state))
           (sample-seq))))]
    (sample-seq)))
\end{lstlisting}
\texttt{anglican.importance/{\linebreak[0]}infer} just calls
\texttt{anglican.inference/{\linebreak[0]}exec} and relies on default
implementations of checkpoint handlers. A different inference
algorithm would provide its own implementations of
\texttt{checkpoint} for \texttt{sample}, \texttt{observe}, or
both, as well as invoke \texttt{exec} from an elaborated
conditional control flow. LMH (\texttt{anglican.lmh}) and SMC
(\texttt{anglican.smc}) are examples of inference algorithms
where either \texttt{observe} (SMC) or \texttt{sample} (LMH)
handler is overridden. In addition, SMC runs multiple particles
(program instances) simultaneously, while LMH re-runs
programs from an intermediate continuation rather than from the
beginning.

>>>>

Implementing an inference algorithm in Anglican amounts
to defining a version of the \texttt{infer} function and 
checkpoint handlers appropriate for the algorithm. The definitions may 
just reuse default implementations or override them
with new algorithm-specific treatment of \texttt{sample} and \texttt{observe} forms 
and inference state.

Let us illustrate this implementation process with
\textit{importance sampling},
the simplest inference algorithm.
Here is an implementation of the \texttt{infer} function for the algorithm:
\begin{lstlisting}[style=default]
(derive ::algorithm
        :anglican.inference/algorithm)
;; invoked when algo parameter is :importance
(defmethod infer :importance 
  [algo prog value & {}]
  (letfn
    ;; recursive function without any parameter
    [sample-seq 
      []
      ;; lazy infinite sequence 
      (lazy-seq
        (cons
          (:state (exec ::algorithm prog value
                        initial-state))
          (sample-seq))))]
    (sample-seq)))
\end{lstlisting}
This implementation is dispatched 
when \texttt{infer} receives \texttt{:importance} as its \texttt{algo} 
parameter. (Anglican includes multiple implementations of \texttt{infer}
for different inference algorithms.) Once dispatched, it lazily 
constructs an infinite sequence of inference states by repeatedly 
executing the program \texttt{prog} using \texttt{exec}
and retrieving the final inference state of the execution 
using \texttt{:state}. For checkpoint handlers, importance sampling simply
relies on their default implementations.
%\texttt{anglican.importance/{\linebreak[0]}infer} just calls
%\texttt{anglican.inference/{\linebreak[0]}exec} and relies on default
%implementations of checkpoint handlers.

Typically, an inference
algorithm has its own implementations of
\texttt{checkpoint} for \texttt{sample}, \texttt{observe}, or
both, as well as invoke \texttt{exec} from an elaborated
conditional control flow. LMH (\texttt{anglican.lmh}) and SMC
(\texttt{anglican.smc}) are examples of inference algorithms
where either \texttt{observe} (SMC) or \texttt{sample} (LMH)
handler is overridden. In addition, SMC runs multiple particles
(threads executing a program) simultaneously, while LMH re-runs
programs from an intermediate continuation rather than from the
beginning.
